Notes on robust solvers for systems of coupled multiscale PDEs

### 2020-02-28 solver/solver.tex
work on A1
update
Updates from Overleaf

@@

Let $\Omega=\Omega^{+}\cup\Omega^{-}$. Introducing a Lagrange multiplier
$p=\epsilon^{-1}\jump{\nabla u \cdot n}\in Q=H^{-1/2}(\Gamma)$
the weak form of \eqref{eq:coupled} reads: Find $u\in V=H^1_{0, \Gamma_D}(\Omega)$,
$\hat{v}\in\hat{V}=H^1(\Gamma)$, $p\in Q$ such that
%
\begin{equation}\label{eq:coupled_weak}
  \begin{aligned}
    &\int_{\Omega}\nabla u\cdot \nabla v\,\mathrm{d}x &\phantom{+\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &+\epsilon\int_{\Gamma} p v\,\mathrm{d}s &= \int_{\Omega}f v\,\mathrm{d}x \quad &\forall v\in V,\\
    &\phantom{\int_{\Omega}\nabla u\cdot \nabla v} &+{\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v} + \hat{u}\hat{v}\,\mathrm{d}x} &-\int_{\Gamma} p\hat{v}\,\mathrm{d}s &= \int_{\Gamma}\hat{f}\hat{v}\,\mathrm{d}s \quad &\forall \hat{v}\in \hat{V},\\
    &\epsilon\int_{\Gamma}u q\,\mathrm{d}s &-\int_{\Gamma}\hat{u}q\,\mathrm{d}s &\phantom{-\int{\Gamma}q\hat{v}} &= \int_{\Gamma}h q\,\mathrm{d}s \quad &\forall q\in Q.
  \end{aligned}
\end{equation}

In the following it will be convenient to work with the operator form of
\eqref{eq:coupled_weak}. In particular the left-hand-side defines an operator
%
\begin{equation}\label{eq:operator2d}
  \mathcal{A} = \begin{pmatrix}
    -\Delta_{\Omega} & & -\epsilon T^{\prime}\\
    & -\Delta_{\Gamma} + I & -I\\
    -\epsilon T & -I & \\
  \end{pmatrix} =
  \begin{pmatrix}
    A_2 &  & B_2'\\
    & A_1 & -B_1'\\
    B_2 & -B_1 & \\
   \end{pmatrix}.
\end{equation}
In \cite{kuchta2016preconditioners} it is shown that $\mathcal{A}$ is an
isomorphism from $W$ to its dual space where
\[
W=H^1(\Omega)_{0, \Gamma_D}\times H^1{(\Gamma)}\times(\epsilon H^{-1/2}(\Gamma)\cap H^{-1}(\Gamma)).
\]
It is further shown that discretization in terms of continuous linear Lagrange
elements (P1) for all the unknowns is inf-sup stable and in turn the diagonal operator
\begin{equation}\label{eq:P1_precond}
\mathcal{B}=\text{diag}(-\Delta_{\Omega}, -\Delta_{\Gamma}+I, \epsilon^2(-\Delta_{\Gamma})^{-1/2}+(-\Delta_{\Gamma}+I)^{-1})^{-1}
\end{equation}
is a robust preconditioner for $\mathcal{A}$. For the sake of completness
(and verifying the implementation) Table \ref{tab:P1} shows the number of
MinRes iterations and condition numbers\footnote{
  Unless specified otherwise all the experiments are done on $\Omega$ a unit square
  triangulated in $2N^2$ isosceles triangles. MinRes iterations are always started
  from a random initial vector and terminate one the relative preconditioned residual
  norm is reduced by factor $10^{8}$. Condition numbers of the preconditioned problem
  are computed by direct solvers, i.e. from the full spectrum.
} of the preconditioned problem for
different levels of refinement and values of the parameter $\epsilon$. The
stability is evident.
%
\begin{table}
    \footnotesize{
      \begin{minipage}{0.49\textwidth}
          \begin{center}
            \begin{tabular}{c|cccc} \hline
              \multirow{2}{*}{$\epsilon$} & \multicolumn{4}{c}{$h$}\\
              \cline{2-5}
 &   $2^{-2}$  & $2^{-3}$ & $2^{-4}$ & $2^{-5}$   \\  
\hline
$10^{6}$ &	7.73&	7.82&	7.86&	7.87\\
$10^{4}$ &	7.73&	7.82&	7.86&	7.87\\
$10^{2}$   &	7.73&	7.82&	7.86&	7.87\\
1     &	7.39&	7.79&	8.01&	8.13\\
$10^{-2}$  &	2.62&	2.63&	2.63&	2.65\\
$10^{-4}$ &	2.62&	2.62&	2.62&	2.62\\
$10^{-6}$ &	2.62&	2.62&	2.62&	2.62\\
\hline
\end{tabular}
\end{center}
      \end{minipage}
    }
          \footnotesize{
            \begin{minipage}{0.49\textwidth}
              \begin{center}
            \begin{tabular}{c|cccccc} \hline
              \multirow{2}{*}{$\epsilon$} & \multicolumn{6}{c}{$h$}\\
              \cline{2-7}
 &   $2^{-3}$ & $2^{-4}$ & $2^{-5}$ & $2^{-6}$ & $2^{-7}$ & $2^{-8}$  \\  
              \hline
$10^{6}$ & 21 & 22 & 22 & 21 & 21 & 19\\
$10^{4}$ & 21 & 22 & 22 & 21 & 21 & 19\\
$10^{2}$ & 21 & 22 & 22 & 21 & 21 & 19\\
1 & 24 & 28 & 26 & 26 & 27 & 25\\
$10^{-2}$ & 10 & 10 & 10 & 12 & 13 & 14\\
$10^{-4}$ & 6 & 6 & 6 & 6 & 7 & 7\\
$10^{-6}$ & 4 & 4 & 4 & 4 & 4 & 4\\
\hline
\end{tabular}                
              \end{center}
        \end{minipage}
      }
          \caption{
            Problem \eqref{eq:coupled_weak} discretized with P1 elements.
            Condition number (left) and iterations counts (right) using
            preconditioner \eqref{eq:P1_precond}.
          }
          \label{tab:P1} 
%\label{table1}
\end{table}

@@

\section{Preconditioning coupled problem}\label{sec:coupled}
\input{coupled}

>>>>

\newtheorem{example}{Example}[section]

@@

Let $\Omega=\Omega^{+}\cup\Omega^{-}$. Introducing a Lagrange multiplier
$p=\epsilon^{-1}\jump{\nabla u \cdot n}\in Q=H^{-1/2}(\Gamma)$
the weak form of \eqref{eq:coupled} reads: Find $u\in V=H^1_{0, \Gamma_D}(\Omega)$,
$\hat{v}\in\hat{V}=H^1(\Gamma)$, $p\in Q$ such that
%
\begin{equation}\label{eq:coupled_weak}
  \begin{aligned}
    &\int_{\Omega}\nabla u\cdot \nabla v\,\mathrm{d}x &\phantom{+\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &+\epsilon\int_{\Gamma} p v\,\mathrm{d}s &= \int_{\Omega}f v\,\mathrm{d}x \quad &\forall v\in V,\\
    &\phantom{\int_{\Omega}\nabla u\cdot \nabla v} &+{\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v} + \hat{u}\hat{v}\,\mathrm{d}s} &-\int_{\Gamma} p\hat{v}\,\mathrm{d}s &= \int_{\Gamma}\hat{f}\hat{v}\,\mathrm{d}s \quad &\forall \hat{v}\in \hat{V},\\
    &\epsilon\int_{\Gamma}u q\,\mathrm{d}s &-\int_{\Gamma}\hat{u}q\,\mathrm{d}s &\phantom{-\int{\Gamma}q\hat{v}} &= \int_{\Gamma}h q\,\mathrm{d}s \quad &\forall q\in Q.
  \end{aligned}
\end{equation}

In the following it will be convenient to work with the operator form of
\eqref{eq:coupled_weak}. In particular the left-hand-side defines an operator
%
\begin{equation}\label{eq:operator2d}
  \mathcal{A} = \begin{pmatrix}
    -\Delta_{\Omega} & & \epsilon T^{\prime}\\
    & -\Delta_{\Gamma} + I & -I\\
    \epsilon T & -I & \\
  \end{pmatrix} =
  \begin{pmatrix}
    A_2 &  & B_2'\\
    & A_1 & -B_1'\\
    B_2 & -B_1 & \\
   \end{pmatrix}.
\end{equation}
In \cite{kuchta2016preconditioners} it is shown that $\mathcal{A}$ is an
isomorphism from $W$ to its dual space where
\[
W=H^1(\Omega)_{0, \Gamma_D}\times H^1{(\Gamma)}\times(\epsilon H^{-1/2}(\Gamma)\cap H^{-1}(\Gamma)).
\]
It is further shown that discretization in terms of continuous linear Lagrange
elements (P1) for all the unknowns is inf-sup stable and in turn the diagonal operator
\begin{equation}\label{eq:P1_precond}
\mathcal{B}=\text{diag}(-\Delta_{\Omega}, -\Delta_{\Gamma}+I, \epsilon^2(-\Delta_{\Gamma})^{-1/2}+(-\Delta_{\Gamma}+I)^{-1})^{-1}
\end{equation}
is a robust preconditioner for $\mathcal{A}$. For the sake of completness
(and verifying the implementation) Table \ref{tab:P1} shows the number of
MinRes iterations and condition numbers\footnote{
  Unless specified otherwise all the experiments are done on $\Omega$ a unit square
  triangulated into $2N^2$ isosceles triangles. MinRes iterations are always started
  from a random initial vector and terminate once the relative preconditioned residual
  norm is reduced by factor $10^{8}$. Condition numbers of the preconditioned problem
  are computed by direct solver, i.e. from the full spectrum.
} of the preconditioned problem for
different levels of refinement and values of the parameter $\epsilon$. The
stability is evident.
%
\begin{table}
    \footnotesize{
      \begin{minipage}{0.49\textwidth}
          \begin{center}
            \begin{tabular}{c|cccc} \hline
              \multirow{2}{*}{$\epsilon$} & \multicolumn{4}{c}{$h$}\\
              \cline{2-5}
 &   $2^{-2}$  & $2^{-3}$ & $2^{-4}$ & $2^{-5}$   \\  
\hline
$10^{6}$ &	7.73&	7.82&	7.86&	7.87\\
$10^{4}$ &	7.73&	7.82&	7.86&	7.87\\
$10^{2}$   &	7.73&	7.82&	7.86&	7.87\\
1     &	7.39&	7.79&	8.01&	8.13\\
$10^{-2}$  &	2.62&	2.63&	2.63&	2.65\\
$10^{-4}$ &	2.62&	2.62&	2.62&	2.62\\
$10^{-6}$ &	2.62&	2.62&	2.62&	2.62\\
\hline
\end{tabular}
\end{center}
      \end{minipage}
    }
          \footnotesize{
            \begin{minipage}{0.49\textwidth}
              \begin{center}
            \begin{tabular}{c|cccccc} \hline
              \multirow{2}{*}{$\epsilon$} & \multicolumn{6}{c}{$h$}\\
              \cline{2-7}
 &   $2^{-3}$ & $2^{-4}$ & $2^{-5}$ & $2^{-6}$ & $2^{-7}$ & $2^{-8}$  \\  
              \hline
$10^{6}$ & 21 & 22 & 22 & 21 & 21 & 19\\
$10^{4}$ & 21 & 22 & 22 & 21 & 21 & 19\\
$10^{2}$ & 21 & 22 & 22 & 21 & 21 & 19\\
1 & 24 & 28 & 26 & 26 & 27 & 25\\
$10^{-2}$ & 10 & 10 & 10 & 12 & 13 & 14\\
$10^{-4}$ & 6 & 6 & 6 & 6 & 7 & 7\\
$10^{-6}$ & 4 & 4 & 4 & 4 & 4 & 4\\
\hline
\end{tabular}                
              \end{center}
        \end{minipage}
      }
          \caption{
            Problem \eqref{eq:coupled_weak} discretized with P1 elements.
            Condition number (left) and iterations counts (right) using
            preconditioner \eqref{eq:P1_precond}.
          }
          \label{tab:P1} 
%\label{table1}
\end{table}

@@

\section{Preconditioning 2$d$ coupled problem}\label{sec:coupled2d}
\input{coupled2d}

\section{Preconditioning 3$d$ coupled problem}\label{sec:coupled2d}
\input{coupled3d}
